### 2021-08-17 paperDraft/paperDraft.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
  10pt,
  dvipsnames,enabledeprecatedfontcommands]{scrartcl}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Nesta attacks study},
  pdfauthor={Patrycja Tempska and Rafal Urbaniak},
  colorlinks=true,
  linkcolor=Maroon,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
%\documentclass{article}

% %packages
 \usepackage{booktabs}

\usepackage{multirow}

\usepackage{graphicx}
\usepackage{longtable}
\usepackage{ragged2e}
\usepackage{etex}
%\usepackage{yfonts}
\usepackage{marvosym}
\usepackage[notextcomp]{kpfonts}
\usepackage{nicefrac}
\newcommand*{\QED}{\hfill \footnotesize {\sc Q.e.d.}}
\usepackage{floatrow}

\usepackage[textsize=footnotesize]{todonotes}
%\linespread{1.5}

\setlength{\parindent}{10pt}
\setlength{\parskip}{1pt}

%language
\usepackage{times}
\usepackage{t1enc}
%\usepackage[utf8x]{inputenc}
%\usepackage[polish]{babel}
%\usepackage{polski}
\usepackage[utf8]{inputenc}
\usepackage{mathptmx}
\usepackage[scaled=0.88]{helvet}

%AMS
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{mathtools}

\usepackage{geometry}
 \geometry{a4paper,left=35mm,top=20mm,}

%environments
\newtheorem{fact}{Fact}

%abbreviations
\newcommand{\ra}{\rangle}
\newcommand{\la}{\langle}
\newcommand{\n}{\neg}
\newcommand{\et}{\wedge}
\newcommand{\jt}{\rightarrow}
\newcommand{\ko}[1]{\forall  #1\,}
\newcommand{\ro}{\leftrightarrow}
\newcommand{\exi}[1]{\exists\, {_{#1}}}
\newcommand{\pr}[1]{\mathsf{P}(#1)}
\newcommand{\cost}{\mathsf{cost}}

\newcommand{\odds}{\mathsf{Odds}}
\newcommand{\ind}{\mathsf{Ind}}
\newcommand{\nf}[2]{\nicefrac{#1\,}{#2}}
\newcommand{\R}[1]{\texttt{#1}}
\newcommand{\prr}[1]{\mbox{$\mathtt{P}_{prior}(#1)$}}
\newcommand{\prp}[1]{\mbox{$\mathtt{P}_{posterior}(#1)$}}

\newtheorem{q}{\color{blue}Question}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}

%technical intermezzo
%---------------------

\newcommand{\intermezzoa}{
	\begin{minipage}[c]{13cm}
	\begin{center}\rule{10cm}{0.4pt}

	\tiny{\sc Optional Content Starts}

	\vspace{-1mm}

	\rule{10cm}{0.4pt}\end{center}
	\end{minipage}\nopagebreak 
	}

\newcommand{\intermezzob}{\nopagebreak 
	\begin{minipage}[c]{13cm}
	\begin{center}\rule{10cm}{0.4pt}

	\tiny{\sc Optional Content Ends}

	\vspace{-1mm}

	\rule{10cm}{0.4pt}\end{center}
	\end{minipage}
	}
%--------------------

\newtheorem*{reply*}{Reply}
\usepackage{enumitem}
\newcommand{\question}[1]{\begin{enumerate}[resume,leftmargin=0cm,labelsep=0cm,align=left]
\item #1
\end{enumerate}}

\usepackage{float}

% \setbeamertemplate{blocks}[rounded][shadow=true]
% \setbeamertemplate{itemize items}[ball]
% \AtBeginPart{}
% \AtBeginSection{}
% \AtBeginSubsection{}
% \AtBeginSubsubsection{}
% \setlength{\emergencystretch}{0em}
% \setlength{\parskip}{0pt}

\usepackage[authoryear]{natbib}

%\bibliographystyle{apalike}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{Nesta attacks study}
\author{Patrycja Tempska and Rafal Urbaniak}
\date{}

\begin{document}
\maketitle

\tableofcontents

\hypertarget{section-abstract}{%
\section{Section Abstract}\label{section-abstract}}

This article describes an experimental intervention study based in a
naturalistic, digital setting (Q\&A forum - Reddit), utilizing a
collective intelligence approach to content moderation and reduction of
the level of verbal aggression among a selected group of Reddit users
who regularly attack other community members. Collective Intelligence in
this sense means exploring the collaboration between human and machine
intelligence to develop solutions to social challenges. Artificial
Intelligence was used to detect verbal aggression (personal attacks) and
notify human volunteers about attacks. Volunteers after receiving
notifications employed interventions based on norm or empathy promotion.
We find that only those who were sanctioned with norms-inducing
interventions had their personal attacks' level significantly decreased.

\vspace{1mm}
\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\DecValTok{2}\SpecialCharTok{+}\DecValTok{2} \CommentTok{\#use this formatting for chunks}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 4
\end{verbatim}

\normalsize

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Although much effort has been made in order to tackle the problem of
verbal aggression and harassment online, looking at various reports and
surveys, it remains a common hindrance for people engaging with social
media in their everyday lives. The situation got exacerbated amidst the
COVID19 pandemic, during which a majority of our social life moved to
cyberspace. During this shift, there was an increase in cyberbullying
attitudes and perpetration (Barlett, Simmers, Roth, \& Gentile (2021)),
90\% increase in public reports of illegal online
content\footnote{\href {https://www.aspistrategist.org.au/australias-esafety-commissioner-targets-abuse-online-as-covid-19-supercharges-cyberbullying/}{https://www.aspistrategist.org.au/australias-esafety-commissioner-targets-abuse-online-as-covid-19-supercharges-cyberbullying/}},
including 114\% increase in non-consensual sharing of intimate images,
30\% increase in cyberbullying, as well as 40\% of increase in adults
reporting online harassment. According to a report conducted by company
L1ght \textbackslash footnote\{\textbackslash href
\{\url{https://l1ght.com/Toxicity_during_coronavirus_Report-L1ght.pdf}\}\{\url{https://l1ght.com/Toxicity_during_coronavirus_Report-L1ght.pdf}\}\},
hate speech directed towards China and the Chinese went up by 900\% on
Twitter. Gaming platforms were in the spotlight as well, with a 40\%
increase in toxicity on Discord.

But alongside the growing need for even more efficient and proactive
moderation, the capacity to execute it did not go hand in hand, forcing
companies and policymakers to rethink the current model of moderation
processes and workforce. Due to the COVID19 restrictions including
social distancing, a lot of those serving the role of moderators had to
be sent
home\footnote{\href {https://qz.com/india/1976450/facebook-covid-19-lockdowns-hurt-content-moderation-algorithms/}{https://qz.com/india/1976450/facebook-covid-19-lockdowns-hurt-content-moderation-algorithms/}}
without the ability to work remotely because of the constraints
affiliated with restrictive non-disclosure agreements (NDA) among
others. Curtailing the moderators' workforce was accompanied by more
agency given to algorithms and AI-based moderation. Those changes, as
argued by Gerrard (2020), can be seen as a serious red flag in terms of
safety for all users on online platforms.

\hypertarget{automated-versus-human-based-moderation}{%
\section{Automated versus Human-based
Moderation}\label{automated-versus-human-based-moderation}}

The hindrances and threats that go along with the Artificial
Intelligence-based methods for moderation have been widely debated, with
the most critical discussions revolving around technology performance
(MacAvaney et al. (2019), Schmidt \& Wiegand (2017)). State-of-the-art
solutions are mostly governed by statistical methods including deep
learning and machine learning (LeCun, Bengio, \& Hinton (2015),
Sejnowski (2020), Jordan \& Mitchell (2015)). Their performance is
inherently tied to the amount of data being fed to the system and the
quality of its annotation. At different stages of the process, from
datasets gathering and preparation, annotation to the training or
algorithms themselves, biases seem to be omnipresent (Binns, Veale, Van
Kleek, \& Shadbolt (2017), Geva, Goldberg, \& Berant (2019) Mehrabi,
Morstatter, Saxena, Lerman, \& Galstyan (2021)). Users' of online
services are also creative in their strategies to circumvent automated
content moderation systems and as shown by Gröndahl, Pajola, Juuti,
Conti, \& Asokan (2018), current techniques are vulnerable to the most
common evasion attacks like word changes (inseting typos and leetspeak),
word-boundary changes (inserting or removing whitespace), or word
appending (appending common or non-hateful words like ``love''). Lack of
generalisability of the models - an ability to perform well on datasets
coming from sources other than the one used for training is a serious
shortcoming as well (Yin \& Zubiaga (2021), Swamy, Jamatia, \& Gambäck
(2019), Rosa et al. (2019)). As shown by Wu, Ribeiro, Heer, \& Weld
(2019), Lipton \& Steinhardt (2019), and Musgrave, Belongie, \& Lim
(2020) in practice, creations of models often lacks thorough error
analysis and legitimate experimental methodology, which can result in
non-reproducibility. This is also connected with a potential lack of
thorough understanding of the limitations of the models and spurious
conclusions being made to a wider public. Specifically, @Lipton \&
Steinhardt (2019) distinguishes four dysfunctional patterns occurring in
the current research paradigm in the industry and academia alike. First,
the inability to draw a clear distinction between speculation and
explanation, with the first one often being disguised as the second.
Second, inability for successful identification of the sources of
empirical gains (whether it was problem formulation, optimization of the
heuristics, data-preprocessing, hyperparameter tuning, or perhaps yet
another aspect). Third, ``mathiness'' - the use obscure language and
often covering weak argumentation with the alluring but often apparent
depth of technical jargon. Last but not least - misuse of language. This
includes suggestive definitions without proper explanation of what they
mean in the context (e.g.~inflating good performance in simple NLP tasks
to human-level understanding), overloading the papers with technical
terminology, or suitcase words (those words that can encompass a variety
of meanings, e.g.~consciousness).

Yet another obstacle in the process is the lack of gold standard in
dataset creation and taxonomies of abusive language being used for
instance in the process of annotating different datasets. Frequently
people obtain data from various sources and do not follow any
universally used instructions when it comes to its annotation, leading
to discrepancies between various datasets being tagged within one domain
(e.g.~hate speech). Lack of expert annotators and proper annotation
criteria and instructions are also widespread, with the common practice
hiring untrained workers from Mechanical Turk or other crowdsourcing
platforms.

Although there are some initiatives developed in response, most notably,
functional tests for Hate Speech Detection Models created by Röttger et
al. (2020), or the Online Safety Data Initiative (OSDI) LINK
\url{https://onlinesafetydata.blog.gov.uk/about-us/}, focused on
projects related to improving access to data, standardizing the
description of online harms, as well as creating tools and benchmarks
for evaluation of technologies focused on safety, much effort must be
made before wider adoption of such solutions comes into force.

At the same time, only automated methods can scan through the massive
amount of content being generated every day on different platforms. On
Facebook, there are more than 3B comments and likes daily
(\url{https://martech.org/facebook-3-2-billion-likes-comments-every-day/}),
500M tweets are sent daily on Twitter
(\url{https://www.oberlo.com/blog/twitter-statistics}), and over 2B
comments made by users of Reddit in 2020
(\url{https://old.reddit.com/r/blog/comments/k967mm/reddit_in_2020/})
which is almost 3M comments made daily. With this amount of content,
it's either impossible or extremely costly to scale the moderation
workforce. One can also have doubts about the ethical aspects of hiring
workers who are often unaware of how this kind of task will affect their
well-being. Being submersed in the cyber-Augean stables takes a toll on
many - as examined by Roberts (2014) \& Roberts (2016), workers who are
hired for such tasks are often low-status and low-wage, isolated, and
asked to keep what they've seen in secret under restrictive NDAs.
Screening through the reported user-generated content is connected with
exposure to violent and deeply disturbing materials, with child
pornography, murders, or suicides as examples of the most extreme cases.
This can lead to serious psychological damage, like depression, or PTSD
(Roberts (2014)). Some of the employees filed a lawsuit against Facebook
and as a result, the company agreed to pay \$52M in compensation for
mental health issues developed during the job
(\url{https://www.theverge.com/2020/5/12/21255870/facebook-content-moderator-settlement-scola-ptsd-mental-health}).
Taking into consideration that Facebook employs 15K moderators
(\url{https://www.forbes.com/sites/johnkoetsier/2020/06/09/300000-facebook-content-moderation-mistakes-daily-report-says/?sh=6c6bdbbc54d0})
and most likely more are needed to keep up with the growing amount of
content, with the parallel considerations about the negative effects of
content moderation on mental health, a collaboration between humans and
machines in this area seems inevitable.

\hypertarget{pro-active-and-reactive-moderation}{%
\section{Pro-active and reactive
moderation}\label{pro-active-and-reactive-moderation}}

There are different approaches when it comes to the moderation of online
content. One can follow the workflow of reactive moderation, which
happens once the content is published. Harmful messages can be either
reported by the users of the platforms or automated methods and then
sent for review to moderators. A set of actions can be then taken
depending on the platform and their community guidelines - on the
content or user level. A harmful message can be deleted, made invisible
to other users, or certain profanities can be altered with special signs
to censor them. Depending on the type and amount of infraction, a
particular user can be warned, muted or shadowbanned, or banned from
further participation in the community for a period of time. The
weakness of the reactive method is that the damage is done. Whoever is
the recipient of the abusive message has the chance to see it and
potentially suffer (Hoff \& Mitchell (2009), Keipi, Näsi, Oksanen, \&
Räsänen (2016), Wwachs2019associations). Yet another weakness connected
with relying solely on human reports is the content that is harmful but
unreported by a receipted or any bystander. Although the exact scale of
unreported content is not known, various self-report studies show that a
lot of children, teens, or even adults do not report cyberbullying or
harassment online (LINKS:
\url{https://www.ctvnews.ca/canada/as-the-pandemic-forces-us-online-lgbtq2s-teens-deal-with-cyberbullying-1.5430945}
; \url{https://www.adl.org/free-to-play-2020} ;
\url{https://www.ditchthelabel.org/wp-content/uploads/2017/05/InGameAbuse.pdf}).

Yet another type of moderation can be distinguished as pro-active or
pre-moderation. In pre-moderation, automated methods are either based on
Artificial Intelligence or other less sophisticated tools
(e.g.~blacklists) and can screen the content before it gets published.
If a type of harmful message gets detected, it can be removed before
reaching the recipient. Due to the aforementioned dubious performance of
state-of-the-art statistical methods, particularly low precision, they
are rarely used autonomously.

Pro-active moderation can be utilized using AI or other methods to
promote socially positive engagement. Instead of or in the combination
with punitive solutions like privileges restriction, one can induce
empathy or community norms with counter-speech. Counter speech as
described by Dangerous Speech Project
(\url{https://dangerousspeech.org/counterspeech/}) is ``any direct
response to hateful or harmful speech which seeks to undermine it.'' As
examined by Munger (2017), counter-speech can be effective in the
reduction of racist tweets (although only in the condition in which a
white male with high followers was approaching another white male). In a
study conducted by Bilewicz et al. (2021), a bot \ldots{}

\hypertarget{collective-intelligence-approach-to-counter-speech}{%
\section{Collective Intelligence Approach to
Counter-speech}\label{collective-intelligence-approach-to-counter-speech}}

\hypertarget{experimental-design}{%
\section{Experimental Design}\label{experimental-design}}

This was a 6-months field experiment in a digital setting conducted on a
popular Q\&A and news forum, Reddit (www.reddit.com). We formed
treatment and control groups based on three main criteria: 1. During the
intervention period, we have expected to have 20 active volunteers at
any given time, each willing to conduct 10 interventions daily. Thus, we
needed approximately 200 attacks daily generated by the treatment
groups. 2. Recruitment of users who regularly attack others was
necessary to measure the effect of interventions at the end. 3.
Recruitment of users who were active during the whole preliminary
monitoring period was necessary to minimize the risk of attrition during
the study. First, we spent 9 weeks for preliminary Reddit's monitoring
period to recruit only those users who generate at least one personal
attack per week and were active during the whole period of the
monitoring. As mentioned, sustained activity was crucial to minimize the
risk users becoming inactive during the course of the study. Next, we
have calculated the daily average number of personal attacks generated
by the group who met the above criteria (which resulted in 357 attacks
per day - 1.94 attacks daily per person on average). Knowing that we
need around 200 attacks daily (just enough so our volunteers can keep up
with the volume) we have randomly selected 390 people for our study
groups (195 per group). The rest (304 people) were selected as our
control group.

Three groups were formed - two treatment groups (195 people in each) and
one control group (304 people). The first treatment group received
counter-speech interventions based on normative influence, while the
second one received interventions based on empathetic influence. Users
in the control group did not receive any intervention.

The duration of the experiment, 6 months, was divided into three
2-months periods. The first two months served as a monitoring period to
properly select groups and establish baselines. The next 2 months served
as treatment period, during which groups received counter-speech
comments from volunteers, in response to personal attacks detected by
the Artificial Intelligence-based system. The last 2-months served as
the post-treatment monitoring period to gather the data needed to
evaluate the effectiveness of interventions.

Normative interventions can refer either to general social norms of
civility and respect, community standards, a particular subreddit rule,
or a descriptive norm among others. E.g. Insulting others is against
Reddit's policy. Hey there, we do not call each other like that here.
Empathetic interventions can refer to the emotional state of the
recipient or the sender of the attack or both. They are designed to
evoke an empathetic response. E.g. There is a human being on the other
side who might be hurt by your words. I see you are frustrated but
remember the human.

Cyberviolence was defined in this experiment as a personal attack - any
kind of verbal harassment, insult, or threat directed against the
interlocutor in a text-based conversation online. Those were detected
using Samurai Labs' cyberviolence detection system.

The following hypotheses were formulated:

H1: If a group of human volunteers notified by an AI-based cyberviolence
detection system about cyberviolence generated by the treatment group
users (cyberviolence will be defined as a personal attack, harassment,
or a threat targeted against an interlocutor) responds with
counter-speech interventions, this will result in a decreased
cyberviolence level for the whole group after the intervention period.

H2: If two groups receive different types of interventions
(empathy-based or normative), then the decrease in cyberviolence will be
larger in the case of normative interventions in comparison to the
empathy-based ones.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\vspace{-3mm}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-barlett2021comparing}{}%
Barlett, C. P., Simmers, M. M., Roth, B., \& Gentile, D. (2021).
Comparing cyberbullying prevalence and process before and during the
COVID-19 pandemic. \emph{The Journal of Social Psychology}, 1--11.

\leavevmode\hypertarget{ref-bilewicz2021artificial}{}%
Bilewicz, M., Tempska, P., Leliwa, G., Dowgiałło, M., Tańska, M.,
Urbaniak, R., \& Wroczyński, M. (2021). Artificial intelligence against
hate: Intervention reducing verbal aggression in the social network
environment. \emph{Aggressive Behavior}, \emph{47}(3), 260--266.

\leavevmode\hypertarget{ref-binns2017like}{}%
Binns, R., Veale, M., Van Kleek, M., \& Shadbolt, N. (2017). Like
trainer, like bot? Inheritance of bias in algorithmic content
moderation. \emph{International Conference on Social Informatics},
405--415. Springer.

\leavevmode\hypertarget{ref-gerrard2020covid19}{}%
Gerrard, Y. (2020). \textless? covid19?\textgreater{} The COVID-19
mental health content moderation conundrum. \emph{Social Media+
Society}, \emph{6}(3), 2056305120948186.

\leavevmode\hypertarget{ref-geva2019we}{}%
Geva, M., Goldberg, Y., \& Berant, J. (2019). Are we modeling the task
or the annotator? An investigation of annotator bias in natural language
understanding datasets. \emph{arXiv Preprint arXiv:1908.07898}.

\leavevmode\hypertarget{ref-grondahl2018all}{}%
Gröndahl, T., Pajola, L., Juuti, M., Conti, M., \& Asokan, N. (2018).
All you need is" love" evading hate speech detection. \emph{Proceedings
of the 11th ACM Workshop on Artificial Intelligence and Security},
2--12.

\leavevmode\hypertarget{ref-hoff2009cyberbullying}{}%
Hoff, D. L., \& Mitchell, S. N. (2009). Cyberbullying: Causes, effects,
and remedies. \emph{Journal of Educational Administration}.

\leavevmode\hypertarget{ref-jordan2015machine}{}%
Jordan, M. I., \& Mitchell, T. M. (2015). Machine learning: Trends,
perspectives, and prospects. \emph{Science}, \emph{349}(6245), 255--260.

\leavevmode\hypertarget{ref-keipi2016online}{}%
Keipi, T., Näsi, M., Oksanen, A., \& Räsänen, P. (2016). \emph{Online
hate and harmful content: Cross-national perspectives}. Taylor \&
Francis.

\leavevmode\hypertarget{ref-lecun2015deep}{}%
LeCun, Y., Bengio, Y., \& Hinton, G. (2015). Deep learning.
\emph{Nature}, \emph{521}(7553), 436--444.

\leavevmode\hypertarget{ref-lipton2019troubling}{}%
Lipton, Z. C., \& Steinhardt, J. (2019). Troubling trends in machine
learning scholarship: Some ML papers suffer from flaws that could
mislead the public and stymie future research. \emph{Queue},
\emph{17}(1), 45--77.

\leavevmode\hypertarget{ref-macavaney2019hate}{}%
MacAvaney, S., Yao, H.-R., Yang, E., Russell, K., Goharian, N., \&
Frieder, O. (2019). Hate speech detection: Challenges and solutions.
\emph{PloS One}, \emph{14}(8), e0221152.

\leavevmode\hypertarget{ref-mehrabi2021survey}{}%
Mehrabi, N., Morstatter, F., Saxena, N., Lerman, K., \& Galstyan, A.
(2021). A survey on bias and fairness in machine learning. \emph{ACM
Computing Surveys (CSUR)}, \emph{54}(6), 1--35.

\leavevmode\hypertarget{ref-munger2017tweetment}{}%
Munger, K. (2017). Tweetment effects on the tweeted: Experimentally
reducing racist harassment. \emph{Political Behavior}, \emph{39}(3),
629--649.

\leavevmode\hypertarget{ref-musgrave2020metric}{}%
Musgrave, K., Belongie, S., \& Lim, S.-N. (2020). A metric learning
reality check. \emph{European Conference on Computer Vision}, 681--699.
Springer.

\leavevmode\hypertarget{ref-roberts2014behind}{}%
Roberts, S. T. (2014). \emph{Behind the screen: The hidden digital labor
of commercial content moderation}. University of Illinois at
Urbana-Champaign.

\leavevmode\hypertarget{ref-roberts2016commercial}{}%
Roberts, S. T. (2016). \emph{Commercial content moderation: Digital
laborers' dirty work}.

\leavevmode\hypertarget{ref-rosa2019automatic}{}%
Rosa, H., Pereira, N., Ribeiro, R., Ferreira, P. C., Carvalho, J. P.,
Oliveira, S., \ldots{} Trancoso, I. (2019). Automatic cyberbullying
detection: A systematic review. \emph{Computers in Human Behavior},
\emph{93}, 333--345.

\leavevmode\hypertarget{ref-rottger2020hatecheck}{}%
Röttger, P., Vidgen, B., Nguyen, D., Waseem, Z., Margetts, H., \&
Pierrehumbert, J. (2020). Hatecheck: Functional tests for hate speech
detection models. \emph{arXiv Preprint arXiv:2012.15606}.

\leavevmode\hypertarget{ref-schmidt2017survey}{}%
Schmidt, A., \& Wiegand, M. (2017). A survey on hate speech detection
using natural language processing. \emph{Proceedings of the Fifth
International Workshop on Natural Language Processing for Social Media},
1--10.

\leavevmode\hypertarget{ref-sejnowski2020unreasonable}{}%
Sejnowski, T. J. (2020). The unreasonable effectiveness of deep learning
in artificial intelligence. \emph{Proceedings of the National Academy of
Sciences}, \emph{117}(48), 30033--30038.

\leavevmode\hypertarget{ref-swamy2019studying}{}%
Swamy, S. D., Jamatia, A., \& Gambäck, B. (2019). Studying
generalisability across abusive language detection datasets.
\emph{Proceedings of the 23rd Conference on Computational Natural
Language Learning (CoNLL)}, 940--950.

\leavevmode\hypertarget{ref-wu2019errudite}{}%
Wu, T., Ribeiro, M. T., Heer, J., \& Weld, D. S. (2019). Errudite:
Scalable, reproducible, and testable error analysis. \emph{Proceedings
of the 57th Annual Meeting of the Association for Computational
Linguistics}, 747--763.

\leavevmode\hypertarget{ref-yin2021towards}{}%
Yin, W., \& Zubiaga, A. (2021). Towards generalisable hate speech
detection: A review on obstacles and solutions. \emph{PeerJ Computer
Science}, \emph{7}, e598.

\end{CSLReferences}

\end{document}