### 2021-09-10 bayesianAnalysis/bayesianReport.tex

\documentclass[10pt,dvipsnames,enabledeprecatedfontcommands]{scrartcl}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{
            pdftitle={Bayesian analysis of the NESTA study of interventions against verbal aggression online},
            pdfauthor={Rafal Urbaniak},
            colorlinks=true,
            linkcolor=Maroon,
            citecolor=Blue,
            urlcolor=blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

%\documentclass{article}

% %packages
 \usepackage{booktabs}

\usepackage{multirow}

\usepackage{graphicx}

\usepackage{longtable}
\usepackage{ragged2e}
\usepackage{etex}
%\usepackage{yfonts}
\usepackage{marvosym}
\usepackage[notextcomp]{kpfonts}
\usepackage{nicefrac}
\newcommand*{\QED}{\hfill \footnotesize {\sc Q.e.d.}}
\usepackage{floatrow}

\usepackage[textsize=footnotesize]{todonotes}
%\linespread{1.5}
\usepackage{pdfpages}
\setlength{\parindent}{10pt}
\setlength{\parskip}{1pt}

%language
\usepackage{times}
\usepackage{t1enc}
%\usepackage[utf8x]{inputenc}
%\usepackage[polish]{babel}
%\usepackage{polski}
\usepackage[utf8]{inputenc}
\usepackage{mathptmx}
\usepackage[scaled=0.88]{helvet}

%AMS
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{mathtools}

\usepackage{geometry}
 \geometry{a4paper,left=35mm,top=20mm,}

%environments
\newtheorem{fact}{Fact}

%abbreviations
\newcommand{\ra}{\rangle}
\newcommand{\la}{\langle}
\newcommand{\n}{\neg}
\newcommand{\et}{\wedge}
\newcommand{\jt}{\rightarrow}
\newcommand{\ko}[1]{\forall  #1\,}
\newcommand{\ro}{\leftrightarrow}
\newcommand{\exi}[1]{\exists\, {_{#1}}}
\newcommand{\pr}[1]{\mathsf{P}(#1)}
\newcommand{\cost}{\mathsf{cost}}

\newcommand{\odds}{\mathsf{Odds}}
\newcommand{\ind}{\mathsf{Ind}}
\newcommand{\nf}[2]{\nicefrac{#1\,}{#2}}
\newcommand{\R}[1]{\texttt{#1}}
\newcommand{\prr}[1]{\mbox{$\mathtt{P}_{prior}(#1)$}}
\newcommand{\prp}[1]{\mbox{$\mathtt{P}_{posterior}(#1)$}}

\newtheorem{q}{\color{blue}Question}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}

%technical intermezzo
%---------------------

\newcommand{\intermezzoa}{
	\begin{minipage}[c]{13cm}
	\begin{center}\rule{10cm}{0.4pt}

	\tiny{\sc Optional Content Starts}

	\vspace{-1mm}

	\rule{10cm}{0.4pt}\end{center}
	\end{minipage}\nopagebreak 
	}

\newcommand{\intermezzob}{\nopagebreak 
	\begin{minipage}[c]{13cm}
	\begin{center}\rule{10cm}{0.4pt}

	\tiny{\sc Optional Content Ends}

	\vspace{-1mm}

	\rule{10cm}{0.4pt}\end{center}
	\end{minipage}
	}
%--------------------

\newtheorem*{reply*}{Reply}
\usepackage{enumitem}
\newcommand{\question}[1]{\begin{enumerate}[resume,leftmargin=0cm,labelsep=0cm,align=left]
\item #1
\end{enumerate}}

\usepackage{float}

% \setbeamertemplate{blocks}[rounded][shadow=true]
% \setbeamertemplate{itemize items}[ball]
% \AtBeginPart{}
% \AtBeginSection{}
% \AtBeginSubsection{}
% \AtBeginSubsubsection{}
% \setlength{\emergencystretch}{0em}
% \setlength{\parskip}{0pt}

\usepackage[authoryear]{natbib}

%\bibliographystyle{apalike}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{Bayesian analysis of the NESTA study of interventions against verbal
aggression online}
\author{Rafal Urbaniak}
\date{}

\begin{document}
\maketitle

\tableofcontents

\section{Exploration}\label{exploration}

Load the dataset and take a look first.

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{summaries <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"datasets/Summaries.csv"}\NormalTok{)}
\KeywordTok{head}\NormalTok{(summaries) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{kable}\NormalTok{( }\StringTok{"latex"}\NormalTok{, }\DataTypeTok{booktabs =}\NormalTok{ T) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options =} \KeywordTok{c}\NormalTok{(}\StringTok{"striped"}\NormalTok{, }\StringTok{"scale_down"}\NormalTok{) ,}\DataTypeTok{font_size =} \DecValTok{9}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}
\centering\begingroup\fontsize{9}{11}\selectfont

\resizebox{\linewidth}{!}{
\begin{tabular}{rlrrrrrrrrrrlr}
\toprule
X & author & AB & AD & AA & CB & CD & CA & Adiff & Cdiff & AdiffS & CdiffS & group & IC\\
\midrule
\cellcolor{gray!6}{1} & \cellcolor{gray!6}{\_swf} & \cellcolor{gray!6}{19} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{720} & \cellcolor{gray!6}{25} & \cellcolor{gray!6}{28} & \cellcolor{gray!6}{-19} & \cellcolor{gray!6}{-692} & \cellcolor{gray!6}{-0.0245122} & \cellcolor{gray!6}{-0.3501491} & \cellcolor{gray!6}{normative} & \cellcolor{gray!6}{1}\\
2 & -Allergic & 24 & 24 & 8 & 1614 & 1451 & 1237 & -16 & -377 & 0.0719197 & 0.1057675 & normative & 3\\
\cellcolor{gray!6}{3} & \cellcolor{gray!6}{-funny-username-} & \cellcolor{gray!6}{23} & \cellcolor{gray!6}{6} & \cellcolor{gray!6}{12} & \cellcolor{gray!6}{847} & \cellcolor{gray!6}{497} & \cellcolor{gray!6}{721} & \cellcolor{gray!6}{-11} & \cellcolor{gray!6}{-126} & \cellcolor{gray!6}{0.2326395} & \cellcolor{gray!6}{0.4690535} & \cellcolor{gray!6}{control} & \cellcolor{gray!6}{0}\\
4 & -Johnny- & 18 & 2 & 8 & 1465 & 408 & 684 & -10 & -781 & 0.2647835 & -0.4789637 & empathy & 2\\
\cellcolor{gray!6}{5} & \cellcolor{gray!6}{1secwhileiyeet3} & \cellcolor{gray!6}{15} & \cellcolor{gray!6}{3} & \cellcolor{gray!6}{4} & \cellcolor{gray!6}{1384} & \cellcolor{gray!6}{198} & \cellcolor{gray!6}{120} & \cellcolor{gray!6}{-11} & \cellcolor{gray!6}{-1264} & \cellcolor{gray!6}{0.2326395} & \cellcolor{gray!6}{-1.1780359} & \cellcolor{gray!6}{control} & \cellcolor{gray!6}{0}\\
\addlinespace
6 & 20CharsIsNotEnough & 16 & 10 & 25 & 779 & 907 & 972 & 9 & 193 & 0.8755188 & 0.9307596 & empathy & 4\\
\bottomrule
\end{tabular}}
\endgroup{}
\end{table}

\normalsize

The basic variables we are dealing with are in the following table.

\begin{table}
\centering\begingroup\fontsize{9}{11}\selectfont

\begin{tabular}{ll}
\toprule
variable & explanation\\
\midrule
\cellcolor{gray!6}{AB} & \cellcolor{gray!6}{attacks before (pre-treatment)}\\
AD & attacks during (the treatment period)\\
\cellcolor{gray!6}{AA} & \cellcolor{gray!6}{attacks after (post-treatment)}\\
CB & comments before\\
\cellcolor{gray!6}{CD} & \cellcolor{gray!6}{comments during}\\
\addlinespace
CA & comments after\\
\cellcolor{gray!6}{group} & \cellcolor{gray!6}{treatment group}\\
IC & intervention count\\
\bottomrule
\end{tabular}
\endgroup{}
\end{table}

Further variables are defined in terms of those, in particular, we will
be predicting \textsf{AdiffS} which is the standardized difference
\textsf{AA}-\textsf{AB}, and \textsf{AdiffS}, which is the standardized
difference \textsf{CA}-\textsf{CB}. Before we proceed, we will also
standardize the predictors, and add a numerical index for the group:

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{summaries}\OperatorTok{$}\NormalTok{ABS <-}\StringTok{ }\KeywordTok{standardize}\NormalTok{(summaries}\OperatorTok{$}\NormalTok{AB)}
\NormalTok{summaries}\OperatorTok{$}\NormalTok{CBS <-}\StringTok{ }\KeywordTok{standardize}\NormalTok{(summaries}\OperatorTok{$}\NormalTok{CB)}
\NormalTok{summaries}\OperatorTok{$}\NormalTok{AAS <-}\StringTok{ }\KeywordTok{standardize}\NormalTok{(summaries}\OperatorTok{$}\NormalTok{AA)}
\NormalTok{summaries}\OperatorTok{$}\NormalTok{CAS <-}\StringTok{ }\KeywordTok{standardize}\NormalTok{(summaries}\OperatorTok{$}\NormalTok{CA)}
\NormalTok{summaries}\OperatorTok{$}\NormalTok{CDS <-}\StringTok{ }\KeywordTok{standardize}\NormalTok{(summaries}\OperatorTok{$}\NormalTok{CD)}
\NormalTok{summaries}\OperatorTok{$}\NormalTok{ADS <-}\StringTok{ }\KeywordTok{standardize}\NormalTok{(summaries}\OperatorTok{$}\NormalTok{AD)}
\NormalTok{summaries}\OperatorTok{$}\NormalTok{group <-}\StringTok{ }\KeywordTok{as.factor}\NormalTok{(summaries}\OperatorTok{$}\NormalTok{group)}
\NormalTok{summaries}\OperatorTok{$}\NormalTok{groupID <-}\StringTok{  }\KeywordTok{as.integer}\NormalTok{( }\KeywordTok{as.factor}\NormalTok{(summaries}\OperatorTok{$}\NormalTok{group) )}
\end{Highlighting}
\end{Shaded}

\normalsize

First, let's take a look at the distribution of \textsf{IC} in the
treatment groups:

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{(summaries[summaries}\OperatorTok{$}\NormalTok{group }\OperatorTok{!=}\StringTok{ "control"}\NormalTok{,], }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ IC, }\DataTypeTok{fill =}\NormalTok{ group))}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_bar}\NormalTok{()}\OperatorTok{+}\KeywordTok{theme_tufte}\NormalTok{()}\OperatorTok{+}
\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{"interventions received"}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"Intervention counts in treatment groups"}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_x_continuous}\NormalTok{(}\DataTypeTok{breaks =} \KeywordTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{40}\NormalTok{,}\DecValTok{5}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{bayesianReport_files/figure-latex/unnamed-chunk-4-1} \end{center}

\normalsize

\todo{Note there were much more empathetic interventions, this needs an explanation}

Second, when we look at the distribution of standardized difference in
attacks, when restricted to (-1,1), the peaks of distributions are
shifted a bit, with lowest median for the normative group, but not too
much:

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{violAdiffS <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(summaries, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{group, }\DataTypeTok{y =}\NormalTok{ AdiffS))}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_violin}\NormalTok{() }\OperatorTok{+}\KeywordTok{theme_tufte}\NormalTok{() }
\NormalTok{violJoint <-}\StringTok{ }\KeywordTok{ggarrange}\NormalTok{(violAdiffS}\OperatorTok{+}\KeywordTok{ggtitle}\NormalTok{(}\StringTok{"whole range"}\NormalTok{),}
\NormalTok{              violAdiffS }\OperatorTok{+}\StringTok{ }\KeywordTok{ylim}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\OperatorTok{-}\DecValTok{1}\NormalTok{,}\DecValTok{1}\NormalTok{))}\OperatorTok{+}\KeywordTok{geom_boxplot}\NormalTok{(}\DataTypeTok{width =} \FloatTok{.2}\NormalTok{)}\OperatorTok{+}
\StringTok{              }\KeywordTok{ggtitle}\NormalTok{(}\StringTok{"restricted to (-1,1)"}\NormalTok{)) }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: Removed 58 rows containing non-finite values (stat_ydensity).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 58 rows containing non-finite values (stat_boxplot).
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{violJointTitled <-}\StringTok{ }\KeywordTok{annotate_figure}\NormalTok{(violJoint, }
  \DataTypeTok{top =} \KeywordTok{text_grob}\NormalTok{(}\StringTok{"Empirical distribution of change in attacks (standardized)"}\NormalTok{,}
                  \DataTypeTok{size =} \DecValTok{12}\NormalTok{))}
\NormalTok{violJointTitled}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{bayesianReport_files/figure-latex/unnamed-chunk-5-1} \end{center}

\normalsize

Analogous plot for comments does not reveal this slight downward shift
for normative, but otherwise the visualisation migth suggest no strong
impact of interventions on attacks, and no impact on comments.

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{violCdiffS <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(summaries, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{group, }\DataTypeTok{y =}\NormalTok{ CdiffS))}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_violin}\NormalTok{() }\OperatorTok{+}\KeywordTok{theme_tufte}\NormalTok{() }
\NormalTok{violJointC <-}\StringTok{ }\KeywordTok{ggarrange}\NormalTok{(violCdiffS}\OperatorTok{+}\KeywordTok{ggtitle}\NormalTok{(}\StringTok{"whole range"}\NormalTok{),}
\NormalTok{              violCdiffS }\OperatorTok{+}\StringTok{ }\KeywordTok{ylim}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\OperatorTok{-}\DecValTok{1}\NormalTok{,}\DecValTok{1}\NormalTok{))}\OperatorTok{+}\KeywordTok{geom_boxplot}\NormalTok{(}\DataTypeTok{width =} \FloatTok{.2}\NormalTok{)}\OperatorTok{+}
\StringTok{              }\KeywordTok{ggtitle}\NormalTok{(}\StringTok{"restricted to (-1,1)"}\NormalTok{)) }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: Removed 90 rows containing non-finite values (stat_ydensity).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 90 rows containing non-finite values (stat_boxplot).
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{violJointCTitled <-}\StringTok{ }\KeywordTok{annotate_figure}\NormalTok{(violJoint, }
  \DataTypeTok{top =} \KeywordTok{text_grob}\NormalTok{(}\StringTok{"Empirical distribution of change in comments (standardized)"}\NormalTok{,}
                  \DataTypeTok{size =} \DecValTok{12}\NormalTok{))}
\NormalTok{violJointCTitled}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{bayesianReport_files/figure-latex/unnamed-chunk-6-1} \end{center}

\normalsize

However, plotting changes against intervention counts reveals that
restricting attention to various activity levels drastically changes the
regression lines.

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{icplot1 <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(summaries, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ IC, }\DataTypeTok{y =}\NormalTok{ AdiffS, }\DataTypeTok{color =}\NormalTok{ group, }\DataTypeTok{fill =}\NormalTok{ group))}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_jitter}\NormalTok{(}\DataTypeTok{alpha =} \FloatTok{0.6}\NormalTok{, }\DataTypeTok{size =}\NormalTok{.}\DecValTok{8}\NormalTok{)}\OperatorTok{+}\KeywordTok{theme_tufte}\NormalTok{()}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_smooth}\NormalTok{(}\DataTypeTok{alpha =} \FloatTok{0.2}\NormalTok{, }\DataTypeTok{method =} \StringTok{"lm"}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{xlim}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{25}\NormalTok{))}\OperatorTok{+}\KeywordTok{ylim}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\OperatorTok{-}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{))}\OperatorTok{+}
\StringTok{  }\KeywordTok{ggtitle}\NormalTok{(}\StringTok{"sd restricted to (-2,2)"}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.65}\NormalTok{, }\FloatTok{0.1}\NormalTok{))}

\NormalTok{icplot2 <-}\StringTok{  }\KeywordTok{ggplot}\NormalTok{(summaries, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ IC, }\DataTypeTok{y =}\NormalTok{ AdiffS, }\DataTypeTok{color =}\NormalTok{ group, }\DataTypeTok{fill =}\NormalTok{ group))}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_jitter}\NormalTok{(}\DataTypeTok{alpha =} \FloatTok{0.6}\NormalTok{, }\DataTypeTok{size =}\NormalTok{.}\DecValTok{8}\NormalTok{)}\OperatorTok{+}\KeywordTok{theme_tufte}\NormalTok{()}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_smooth}\NormalTok{(}\DataTypeTok{alpha =} \FloatTok{0.2}\NormalTok{, }\DataTypeTok{method =} \StringTok{"lm"}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{xlim}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{25}\NormalTok{))}\OperatorTok{+}\KeywordTok{ylim}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\OperatorTok{-}\DecValTok{1}\NormalTok{,}\DecValTok{1}\NormalTok{))}\OperatorTok{+}\KeywordTok{ggtitle}\NormalTok{(}\StringTok{"sd restricted to (-1,1)"}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.65}\NormalTok{, }\FloatTok{0.1}\NormalTok{))}

\NormalTok{icplotJoint <-}\StringTok{ }\KeywordTok{ggarrange}\NormalTok{(icplot1, icplot2) }
\NormalTok{icplotTitled <-}\StringTok{ }\KeywordTok{annotate_figure}\NormalTok{(icplotJoint, }
  \DataTypeTok{top =} \KeywordTok{text_grob}\NormalTok{(}\StringTok{"Change in attacks (standardized) vs interventions received"}\NormalTok{,  }\DataTypeTok{size =} \DecValTok{12}\NormalTok{))}
\NormalTok{icplotTitled}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{bayesianReport_files/figure-latex/unnamed-chunk-7-1} \end{center}

\normalsize

Some interactions are also suggested by the differences in linear
smoothing when attention is restricted when it comes to change in
comments.

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{icCplot1 <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(summaries, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ IC, }\DataTypeTok{y =}\NormalTok{ CdiffS, }\DataTypeTok{color =}\NormalTok{ group, }\DataTypeTok{fill =}\NormalTok{ group))}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_jitter}\NormalTok{(}\DataTypeTok{alpha =} \FloatTok{0.6}\NormalTok{, }\DataTypeTok{size =}\NormalTok{.}\DecValTok{8}\NormalTok{)}\OperatorTok{+}\KeywordTok{theme_tufte}\NormalTok{()}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_smooth}\NormalTok{(}\DataTypeTok{alpha =} \FloatTok{0.2}\NormalTok{, }\DataTypeTok{method =} \StringTok{"lm"}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{xlim}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{25}\NormalTok{))}\OperatorTok{+}\KeywordTok{ylim}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\OperatorTok{-}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{))}\OperatorTok{+}
\StringTok{  }\KeywordTok{ggtitle}\NormalTok{(}\StringTok{"sd restricted to (-2,2)"}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.65}\NormalTok{, }\FloatTok{0.1}\NormalTok{))}

\NormalTok{icCplot2 <-}\StringTok{  }\KeywordTok{ggplot}\NormalTok{(summaries, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ IC, }\DataTypeTok{y =}\NormalTok{ CdiffS, }\DataTypeTok{color =}\NormalTok{ group, }\DataTypeTok{fill =}\NormalTok{ group))}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_jitter}\NormalTok{(}\DataTypeTok{alpha =} \FloatTok{0.6}\NormalTok{, }\DataTypeTok{size =}\NormalTok{.}\DecValTok{8}\NormalTok{)}\OperatorTok{+}\KeywordTok{theme_tufte}\NormalTok{()}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_smooth}\NormalTok{(}\DataTypeTok{alpha =} \FloatTok{0.2}\NormalTok{, }\DataTypeTok{method =} \StringTok{"lm"}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{xlim}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{25}\NormalTok{))}\OperatorTok{+}\KeywordTok{ylim}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\OperatorTok{-}\DecValTok{1}\NormalTok{,}\DecValTok{1}\NormalTok{))}\OperatorTok{+}\KeywordTok{ggtitle}\NormalTok{(}\StringTok{"sd restricted to (-1,1)"}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.65}\NormalTok{, }\FloatTok{0.1}\NormalTok{))}

\NormalTok{icCplotJoint <-}\StringTok{ }\KeywordTok{ggarrange}\NormalTok{(icCplot1, icCplot2) }
\NormalTok{icCplotTitled <-}\StringTok{ }\KeywordTok{annotate_figure}\NormalTok{(icCplotJoint, }
  \DataTypeTok{top =} \KeywordTok{text_grob}\NormalTok{(}\StringTok{"Change in comments (standardized) vs interventions received"}\NormalTok{,}
   \DataTypeTok{size =} \DecValTok{12}\NormalTok{))}
\NormalTok{icCplotTitled}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{bayesianReport_files/figure-latex/unnamed-chunk-8-1} \end{center}

\normalsize

This suggests we should keep an eye out for interactions in the
analysis, and that the intial comparison of means or medians between
groups might be misleading if the effects in different volume groups are
different and cancel each other.

Now, let's inspect correlations between the variables involved in the
model:

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{summariesCorr <-}\StringTok{ }\KeywordTok{select}\NormalTok{(summaries, IC, ABS, CBS, AAS, CAS, CDS, ADS)}
\KeywordTok{ggcorr}\NormalTok{(summariesCorr, }\DataTypeTok{method =} \KeywordTok{c}\NormalTok{(}\StringTok{"pairwise"}\NormalTok{),}
       \DataTypeTok{digits =} \DecValTok{4}\NormalTok{, }\DataTypeTok{low =} \StringTok{"steelblue"}\NormalTok{, }\DataTypeTok{mid =} \StringTok{"white"}\NormalTok{,}
       \DataTypeTok{high =} \StringTok{"darkred"}\NormalTok{, }\DataTypeTok{midpoint =}\DecValTok{0}\NormalTok{,}
       \DataTypeTok{geom =} \StringTok{"tile"}\NormalTok{, }\DataTypeTok{label =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{label_size=}\DecValTok{4}\NormalTok{, }\DataTypeTok{label_round =}\DecValTok{2}\NormalTok{, }\DataTypeTok{layout.exp =}\DecValTok{1}\NormalTok{,}
       \DataTypeTok{label_alpha =} \OtherTok{FALSE}\NormalTok{,}\DataTypeTok{hjust =} \FloatTok{0.75}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{bayesianReport_files/figure-latex/unnamed-chunk-9-1} \end{center}

\normalsize

This tells us that almost no predictors are strongly correlated, except
for pairs \textsf{CBS}-\textsf{CDS}, so we drop CDS from the analysis
and avoid using them in the same model to avoid multicolinearity issues.
These are just comments during the intervention period, which,
unsurprisingly are also a good proxy for comments before and comments
after.

\section{Causal inference}\label{causal-inference}

To identify the right variables to condition (or not condition) on to
identify the causal effect of the interventions, we first need to think
about the causal structure of the problem. Here's a plausible causal
structure that we will be working with:

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{dag <-}\StringTok{ }\KeywordTok{dagitty}\NormalTok{(}\StringTok{"}
\StringTok{  dag\{}
\StringTok{  CDS -> ADS -> IC  ons}
\StringTok{               U [unobserved]   }
\StringTok{               U -> CBS -> ABS  }
\StringTok{               U -> ABS        }
\StringTok{               U -> CDS -> ADS  }
\StringTok{               U -> ADS         }
\StringTok{               U -> CAS -> AAS    }
\StringTok{               U -> AAS                        }
\StringTok{               IC -> AAS        }
\StringTok{               IC -> CAS        }
\StringTok{               IT -> CAS        }
\StringTok{               IT -> AAS}
\StringTok{               CBS -> CDS -> CAS}
\StringTok{               ABS -> ADS -> AAS}
\StringTok{               \}"}\NormalTok{)}
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{123}\NormalTok{)}
\KeywordTok{drawdag}\NormalTok{(dag)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{bayesianReport_files/figure-latex/unnamed-chunk-10-1} \end{center}

\normalsize

Comments during impact attacks during, which trigger interventions.
Unmeasured user features cause comments before, which impact attacks
before, and also attacks before directly. Comments during (their impact
on ADS is areadly included) impact attacks during during directly and
comments after, which impact attacks after and attacks after directly.
Intervention count impacts attacks after and comments after. The same
directions of impact are included for intervention type. Finally,
comments through time are connected causally, and so are attacks.

We already know not to condition on CDS if we condition on CAS or CBS.
What else? \textsf{IT} has no bacwkard paths, but \textsf{IC} does.
Let's identify all paths from \textsf{IC} to \textsf{AAS}:

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{paths}\NormalTok{(dag, }\DataTypeTok{from =} \KeywordTok{c}\NormalTok{(}\StringTok{"IC"}\NormalTok{), }\DataTypeTok{to =} \StringTok{"AAS"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## $paths
##  [1] "IC -> AAS"                                              
##  [2] "IC -> CAS -> AAS"                                       
##  [3] "IC -> CAS <- CDS -> ADS -> AAS"                         
##  [4] "IC -> CAS <- CDS -> ADS <- ABS <- CBS <- U -> AAS"      
##  [5] "IC -> CAS <- CDS -> ADS <- ABS <- U -> AAS"             
##  [6] "IC -> CAS <- CDS -> ADS <- U -> AAS"                    
##  [7] "IC -> CAS <- CDS <- CBS -> ABS -> ADS -> AAS"           
##  [8] "IC -> CAS <- CDS <- CBS -> ABS -> ADS <- U -> AAS"      
##  [9] "IC -> CAS <- CDS <- CBS -> ABS <- U -> AAS"             
## [10] "IC -> CAS <- CDS <- CBS -> ABS <- U -> ADS -> AAS"      
## [11] "IC -> CAS <- CDS <- CBS <- U -> AAS"                    
## [12] "IC -> CAS <- CDS <- CBS <- U -> ABS -> ADS -> AAS"      
## [13] "IC -> CAS <- CDS <- CBS <- U -> ADS -> AAS"             
## [14] "IC -> CAS <- CDS <- U -> AAS"                           
## [15] "IC -> CAS <- CDS <- U -> ABS -> ADS -> AAS"             
## [16] "IC -> CAS <- CDS <- U -> ADS -> AAS"                    
## [17] "IC -> CAS <- CDS <- U -> CBS -> ABS -> ADS -> AAS"      
## [18] "IC -> CAS <- IT -> AAS"                                 
## [19] "IC -> CAS <- U -> AAS"                                  
## [20] "IC -> CAS <- U -> ABS -> ADS -> AAS"                    
## [21] "IC -> CAS <- U -> ABS <- CBS -> CDS -> ADS -> AAS"      
## [22] "IC -> CAS <- U -> ADS -> AAS"                           
## [23] "IC -> CAS <- U -> CBS -> ABS -> ADS -> AAS"             
## [24] "IC -> CAS <- U -> CBS -> CDS -> ADS -> AAS"             
## [25] "IC -> CAS <- U -> CDS -> ADS -> AAS"                    
## [26] "IC -> CAS <- U -> CDS <- CBS -> ABS -> ADS -> AAS"      
## [27] "IC <- ADS -> AAS"                                       
## [28] "IC <- ADS <- ABS <- CBS -> CDS -> CAS -> AAS"           
## [29] "IC <- ADS <- ABS <- CBS -> CDS -> CAS <- IT -> AAS"     
## [30] "IC <- ADS <- ABS <- CBS -> CDS -> CAS <- U -> AAS"      
## [31] "IC <- ADS <- ABS <- CBS -> CDS <- U -> AAS"             
## [32] "IC <- ADS <- ABS <- CBS -> CDS <- U -> CAS -> AAS"      
## [33] "IC <- ADS <- ABS <- CBS -> CDS <- U -> CAS <- IT -> AAS"
## [34] "IC <- ADS <- ABS <- CBS <- U -> AAS"                    
## [35] "IC <- ADS <- ABS <- CBS <- U -> CAS -> AAS"             
## [36] "IC <- ADS <- ABS <- CBS <- U -> CAS <- IT -> AAS"       
## [37] "IC <- ADS <- ABS <- CBS <- U -> CDS -> CAS -> AAS"      
## [38] "IC <- ADS <- ABS <- CBS <- U -> CDS -> CAS <- IT -> AAS"
## [39] "IC <- ADS <- ABS <- U -> AAS"                           
## [40] "IC <- ADS <- ABS <- U -> CAS -> AAS"                    
## [41] "IC <- ADS <- ABS <- U -> CAS <- IT -> AAS"              
## [42] "IC <- ADS <- ABS <- U -> CBS -> CDS -> CAS -> AAS"      
## [43] "IC <- ADS <- ABS <- U -> CBS -> CDS -> CAS <- IT -> AAS"
## [44] "IC <- ADS <- ABS <- U -> CDS -> CAS -> AAS"             
## [45] "IC <- ADS <- ABS <- U -> CDS -> CAS <- IT -> AAS"       
## [46] "IC <- ADS <- CDS -> CAS -> AAS"                         
## [47] "IC <- ADS <- CDS -> CAS <- IT -> AAS"                   
## [48] "IC <- ADS <- CDS -> CAS <- U -> AAS"                    
## [49] "IC <- ADS <- CDS <- CBS -> ABS <- U -> AAS"             
## [50] "IC <- ADS <- CDS <- CBS -> ABS <- U -> CAS -> AAS"      
## [51] "IC <- ADS <- CDS <- CBS -> ABS <- U -> CAS <- IT -> AAS"
## [52] "IC <- ADS <- CDS <- CBS <- U -> AAS"                    
## [53] "IC <- ADS <- CDS <- CBS <- U -> CAS -> AAS"             
## [54] "IC <- ADS <- CDS <- CBS <- U -> CAS <- IT -> AAS"       
## [55] "IC <- ADS <- CDS <- U -> AAS"                           
## [56] "IC <- ADS <- CDS <- U -> CAS -> AAS"                    
## [57] "IC <- ADS <- CDS <- U -> CAS <- IT -> AAS"              
## [58] "IC <- ADS <- U -> AAS"                                  
## [59] "IC <- ADS <- U -> ABS <- CBS -> CDS -> CAS -> AAS"      
## [60] "IC <- ADS <- U -> ABS <- CBS -> CDS -> CAS <- IT -> AAS"
## [61] "IC <- ADS <- U -> CAS -> AAS"                           
## [62] "IC <- ADS <- U -> CAS <- IT -> AAS"                     
## [63] "IC <- ADS <- U -> CBS -> CDS -> CAS -> AAS"             
## [64] "IC <- ADS <- U -> CBS -> CDS -> CAS <- IT -> AAS"       
## [65] "IC <- ADS <- U -> CDS -> CAS -> AAS"                    
## [66] "IC <- ADS <- U -> CDS -> CAS <- IT -> AAS"              
## 
## $open
##  [1]  TRUE  TRUE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE
## [13] FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE
## [25] FALSE FALSE  TRUE  TRUE FALSE FALSE FALSE FALSE FALSE  TRUE  TRUE FALSE
## [37]  TRUE FALSE  TRUE  TRUE FALSE  TRUE FALSE  TRUE FALSE  TRUE FALSE FALSE
## [49] FALSE FALSE FALSE  TRUE  TRUE FALSE  TRUE  TRUE FALSE  TRUE FALSE FALSE
## [61]  TRUE FALSE  TRUE FALSE  TRUE FALSE
\end{verbatim}

\normalsize

Crucially, all backdoor paths go through \textsf{ADS}, which then
becomes either a fork or a pipe, so all backdoor paths can be closed by
conditioning on \textsf{ADS}. Moreover there is only one directed
indirect path, it goes through \textsf{CAS}, so we should not condition
on it if we are to identify causal effect on attacks mediated by impact
on comments (unless we care about the direct effect of \textsf{IC} and
\textsf{IT} on \textsf{AAS}, but that's a separate question). This is in
line with the adjustment set identified algorithmically, and the same
move makes sense when we want to predict \textsf{CAS}.

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{adjustmentSets}\NormalTok{(dag, }\DataTypeTok{exposure =} \KeywordTok{c}\NormalTok{(}\StringTok{"IC"}\NormalTok{, }\StringTok{"IT"}\NormalTok{), }\DataTypeTok{outcome =} \StringTok{"AAS"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## { ADS }
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{adjustmentSets}\NormalTok{(dag, }\DataTypeTok{exposure =} \KeywordTok{c}\NormalTok{(}\StringTok{"IC"}\NormalTok{, }\StringTok{"IT"}\NormalTok{), }\DataTypeTok{outcome =} \StringTok{"CAS"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## { ADS }
\end{verbatim}

\normalsize

It's open season for other variables, and our decision to include them
in the model will be guided by information-theoretic criteria of
predictive power.

\section{Bayesian models and their
priors}\label{bayesian-models-and-their-priors}

We will focus on a class of additive models where the outcome variable
is normally distributed around the predicted mean, which is a linear
function of predictors (possibly with some interactions). To spoil the
story, we will end up using a model, whose specification is as follows:

\begin{align*}
\mathsf{AdiffS} & \sim Norm(\mu, \sigma)\\
mu_i & = \alpha + \beta_{\mathsf{ADS}}[\mathsf{ADS}] + \beta_{\mathsf{IT}}[\mathsf{group}] +
 \beta_{IC}[\mathsf{group}]\times \mathsf{IC} + b_{\mathsf{ADSIC}}\times \mathsf{ADS} \times \mathsf{IC} + \beta_{\mathsf{CBS}}[\mathsf{group}] \times \mathsf{CBS}
\end{align*}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\vspace{-3mm}

\end{document}