### 2021-12-21 technicalReport/technicalReport2.Rmd

---
title: "Bayesian analysis of the NESTA study of interventions against  verbal aggression online \\linebreak  Technical Report"
author: "Rafal Urbaniak"
output:
  pdf_document:
  number_sections: yes
df_print: kable
keep_tex: yes
includes:
  in_header: Rafal_latex6.sty
html_document:
  df_print: paged
word_document: default
classoption: dvipsnames, enabledeprecatedfontcommands
fontsize: 10pt
documentclass: scrartcl
urlcolor: blue
bibliography: ../references/attacks.bib
csl: ../references/apa-6th-edition.csl
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '../')

#libraries used
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(kableExtra)
library(viridis)
library(rethinking)
library(ggplot2)
library(ggpubr)
library(tidyverse)
library(GGally)
library(dagitty)
library(reshape)
library(lubridate)
library(formatR)

mykable <- function(object) {kable(object, "latex", booktabs = T) %>% kable_styling(latex_options = "striped",font_size = 9)}

removeX <-   theme(axis.title.x=element_blank(),
                   axis.text.x=element_blank(),
                   axis.ticks.x=element_blank())

removeY <-   theme(axis.title.y=element_blank(),
                   axis.text.y=element_blank(),
                   axis.ticks.y=element_blank())

knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 70), tidy = TRUE)

#kable(table(data$group), "latex", booktabs = T, col.names = c("Group", "n")) %>% 
#  kable_styling(latex_options = c("striped","HOLD_position"),font_size = 9) 
```

\tableofcontents

# Data and exploration

\vspace{1mm}
\scriptsize

```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE, tidy = TRUE}
Hate <- readRDS(file = "datasets/RAWNESTA/Hate.rds")
Comments <-   readRDS(file = "datasets/RAWNESTA/Comments.rds")
summaries <- read.csv(file = "datasets/Summaries.csv")

dates <- colnames(Hate)[-1]
dates <- as.Date(dates)
startDate <- dates[1]
interventionDate <- "2020-07-08"
observationDate <- "2020-09-09"
end <- dates[length(dates)]

periods<- numeric(length(dates))
periods <- ifelse(dates < interventionDate,"pre-treatment",periods)
periods <- ifelse(dates >= interventionDate & dates < observationDate,"treatment",periods)
periods <- ifelse(dates >= observationDate,"post-treatment",periods)

hateTS <- as.data.frame(colSums(Hate[,-1]))
hateTS$date <- as.Date(rownames(hateTS))
rownames(hateTS) <- NULL
colnames(hateTS) <- c("attacks","date")
hateTS$periods <- periods

interventions <- readRDS(file = "datasets/interventions.rds")
interventionsTS <- as.data.frame(table(interventions$day))
interventionsTS$Var1 <- as.Date(interventionsTS$Var1)
colnames(interventionsTS) <- c("date", "interventions")

periodsDF <- merge(x = hateTS, y = interventionsTS, by = "date", all.x = TRUE)

idx <- c(1, diff(periodsDF$date))
i2 <- c(1,which(idx != 1), nrow(periodsDF)+1)
periodsDF$grp <- rep(1:length(diff(i2)), diff(i2))

periodsDF$interventions[is.na(periodsDF$interventions) & periodsDF$periods == "treatment"] <- 0

periodsPlot <- ggplot(periodsDF)+
  geom_line(aes(x=date, y = attacks, group = grp),
             alpha = 0.8, size = .6)+
  geom_line(aes(x=date, y = interventions, group = grp),
            alpha = 0.8, size = .6)+
  geom_vline(xintercept = startDate, lty =2, size =.2, alpha=0.5)+
  geom_vline(xintercept = as.Date(interventionDate), lty =2, size =.2, alpha=0.5)+
  geom_vline(xintercept =  as.Date(observationDate), lty = 2, size =.2, alpha=0.5 )+
  geom_vline(xintercept = as.Date(end), lty =2, size =.2, alpha=0.5)+
  labs(title = "Attacks and interventions time series",
       subtitle = "no line at data gaps", 
       caption = "days with data: 81 (pre-treatment), 62 (treatment), 72 (post-treatment)")+
  theme_tufte() + theme(axis.title.x=element_blank(), 
      plot.caption = element_text(hjust = 0.5,  face= "italic"))+
  scale_x_date(date_labels = "%b %d", breaks = c(startDate, as.Date(startDate), as.Date(interventionDate), as.Date(observationDate), end), 
               limits = c(startDate-30,end+10))+ylab("count")+
  annotate("rect", xmin = as.Date(interventionDate), xmax = as.Date(observationDate), ymin = -1, ymax = 360,
           alpha = .2,fill = "darkgreen")+ylim(c(-1,370))+
  annotate("text", label = "pre-treatment", x = as.Date(startDate)+2, y = 370, hjust =0 )+
  annotate("text", label = "treatment", x = as.Date(interventionDate)+2, y = 370, hjust =0 )+
  annotate("text", label = "post-treatment", x = as.Date(observationDate)+2, y = 370, hjust =0 )+
  annotate("text", label = "interventions:", x = as.Date(interventionDate)-52, y = 15, hjust =0 )+
  annotate("text", label = "attacks:", x = as.Date(startDate)-30, y = 215, hjust =0 )

periodsDF$weekdays <-  weekdays(as.Date(periodsDF$date))
periodsDF$weeks <-  week(as.Date(periodsDF$date))

periodsDF$weekdays <- as.factor(periodsDF$weekdays)
levels(periodsDF$weekdays) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday",
                                "Saturday", "Sunday")
weeksPlot <- ggplot(periodsDF)+
  geom_smooth(aes(x = weekdays, y = attacks, group =1))+
  geom_line(aes(x = weekdays, y = attacks, group = weeks),alpha = 0.1)+
  theme_tufte()+labs(title="Weekly attacks over six months", subtitle = "No weekly patterns")+xlab("")+
  ylab("count")
```
\normalsize

For the duration of the project we selected `r length(unique(summaries$author))` Reddit users and tracked their activity, starting on `r `startDate`, beginning the intervention period on `r interventionDate`, leading to a further observation period starting on  `r observationDate` and ending on  `r end`. The time series of attacks observed and of interventions conducted can be inspected in Figure \ref{fig:periodsPlot}.

\begin{figure}
```{r periodsPlot,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE}
periodsPlot
```
\caption{Daily sums of attacks and interventions throughout the three experimental periods.}
\label{fig:periodsPlot}
\end{figure}

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
#head(summaries) %>% kable( "latex", booktabs = T) %>% 
#  kable_styling(latex_options = c("striped", "scale_down") ,font_size = 9)

```
\normalsize

The basic variables we are dealing with are in the following table.

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
variable <- c("AB", "AD", "AA", "CB", "CD", "CA", "group", "IC")
explanation <- c("attacks before (pre-treatment)", "attacks during (the treatment period)",
                 "attacks after (post-treatment)", "comments before", "comments during",
                 "comments after", "treatment group", "intervention count")
vars <- data.frame(variable, explanation)
mykable(vars)
```