### 2021-12-21 technicalReport/technicalReport3.Rmd

---
title: "Bayesian analysis of the NESTA study of interventions against  verbal aggression online \\linebreak  Technical Report"
author: "Rafal Urbaniak"
output:
  pdf_document:
  number_sections: yes
df_print: kable
keep_tex: yes
includes:
  in_header: Rafal_latex7.sty
html_document:
  df_print: paged
word_document: default
classoption: dvipsnames, enabledeprecatedfontcommands
fontsize: 10pt
documentclass: scrartcl
urlcolor: blue
bibliography: ../references/attacks.bib
csl: ../references/apa-6th-edition.csl
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '../')

#libraries used
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(kableExtra)
library(viridis)
library(rethinking)
library(ggplot2)
library(ggpubr)
library(tidyverse)
library(GGally)
library(dagitty)
library(reshape)
library(lubridate)
library(formatR)

mykable <- function(object) {kable(object, "latex", booktabs = T,linesep = "") %>% kable_styling(latex_options = "striped",font_size = 9)}

removeX <-   theme(axis.title.x=element_blank(),
                   axis.text.x=element_blank(),
                   axis.ticks.x=element_blank())

removeY <-   theme(axis.title.y=element_blank(),
                   axis.text.y=element_blank(),
                   axis.ticks.y=element_blank())

knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 70), tidy = TRUE)

#kable(table(data$group), "latex", booktabs = T, col.names = c("Group", "n")) %>% 
#  kable_styling(latex_options = c("striped","HOLD_position"),font_size = 9) 
```

\tableofcontents

# Data and exploration

\vspace{1mm}
\scriptsize

```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE, tidy = TRUE}
Hate <- readRDS(file = "datasets/RAWNESTA/Hate.rds")
Comments <-   readRDS(file = "datasets/RAWNESTA/Comments.rds")
summaries <- read.csv(file = "datasets/Summaries.csv")

dates <- colnames(Hate)[-1]
dates <- as.Date(dates)
startDate <- dates[1]
interventionDate <- "2020-07-08"
observationDate <- "2020-09-09"
end <- dates[length(dates)]

periods<- numeric(length(dates))
periods <- ifelse(dates < interventionDate,"pre-treatment",periods)
periods <- ifelse(dates >= interventionDate & dates < observationDate,"treatment",periods)
periods <- ifelse(dates >= observationDate,"post-treatment",periods)

hateTS <- as.data.frame(colSums(Hate[,-1]))
hateTS$date <- as.Date(rownames(hateTS))
rownames(hateTS) <- NULL
colnames(hateTS) <- c("attacks","date")
hateTS$periods <- periods

interventions <- readRDS(file = "datasets/interventions.rds")
interventionsTS <- as.data.frame(table(interventions$day))
interventionsTS$Var1 <- as.Date(interventionsTS$Var1)
colnames(interventionsTS) <- c("date", "interventions")

periodsDF <- merge(x = hateTS, y = interventionsTS, by = "date", all.x = TRUE)

idx <- c(1, diff(periodsDF$date))
i2 <- c(1,which(idx != 1), nrow(periodsDF)+1)
periodsDF$grp <- rep(1:length(diff(i2)), diff(i2))

periodsDF$interventions[is.na(periodsDF$interventions) & periodsDF$periods == "treatment"] <- 0

periodsPlot <- ggplot(periodsDF)+
  geom_line(aes(x=date, y = attacks, group = grp),
             alpha = 0.8, size = .6)+
  geom_line(aes(x=date, y = interventions, group = grp),
            alpha = 0.8, size = .6)+
  geom_vline(xintercept = startDate, lty =2, size =.2, alpha=0.5)+
  geom_vline(xintercept = as.Date(interventionDate), lty =2, size =.2, alpha=0.5)+
  geom_vline(xintercept =  as.Date(observationDate), lty = 2, size =.2, alpha=0.5 )+
  geom_vline(xintercept = as.Date(end), lty =2, size =.2, alpha=0.5)+
  labs(title = "Attacks and interventions time series",
       subtitle = "no line at data gaps", 
       caption = "days with data: 81 (pre-treatment), 62 (treatment), 72 (post-treatment)")+
  theme_tufte() + theme(axis.title.x=element_blank(), 
      plot.caption = element_text(hjust = 0.5,  face= "italic"))+
  scale_x_date(date_labels = "%b %d", breaks = c(startDate, as.Date(startDate), as.Date(interventionDate), as.Date(observationDate), end), 
               limits = c(startDate-30,end+10))+ylab("count")+
  annotate("rect", xmin = as.Date(interventionDate), xmax = as.Date(observationDate), ymin = -1, ymax = 360,
           alpha = .2,fill = "darkgreen")+ylim(c(-1,370))+
  annotate("text", label = "pre-treatment", x = as.Date(startDate)+2, y = 370, hjust =0 )+
  annotate("text", label = "treatment", x = as.Date(interventionDate)+2, y = 370, hjust =0 )+
  annotate("text", label = "post-treatment", x = as.Date(observationDate)+2, y = 370, hjust =0 )+
  annotate("text", label = "interventions:", x = as.Date(interventionDate)-55, y = 15, hjust =0 )+
  annotate("text", label = "attacks:", x = as.Date(startDate)-30, y = 215, hjust =0 )

periodsDF$weekdays <-  weekdays(as.Date(periodsDF$date))
periodsDF$weeks <-  week(as.Date(periodsDF$date))

periodsDF$weekdays <- as.factor(periodsDF$weekdays)
levels(periodsDF$weekdays) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday",
                                "Saturday", "Sunday")
weeksPlot <- ggplot(periodsDF)+
  geom_smooth(aes(x = weekdays, y = attacks, group =1))+
  geom_line(aes(x = weekdays, y = attacks, group = weeks),alpha = 0.1)+
  theme_tufte()+labs(title="Personal attacks throught weekdays (six months)", subtitle = "No weekly patterns")+xlab("")+
  ylab("count")
```
\normalsize

For the duration of the project we selected `r length(unique(summaries$author))` Reddit users and tracked their activity (with some breaks resulting from API restrictions and technical issues, which were mostly sorted out in the observation period), starting on `r `startDate`, beginning the intervention period on `r interventionDate`, leading to a further observation period starting on  `r observationDate` and ending on  `r end`. The time series of attacks observed and of interventions conducted can be inspected in Figure \ref{fig:periodsPlot}.

\begin{figure}[H]
```{r periodsPlot,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE}
periodsPlot
```
\caption{Daily sums of attacks and interventions throughout the three experimental periods, with GAM smoothing (blue).}
\label{fig:periodsPlot}
\end{figure}

Interestingly, no weekly patterns of overall aggressive behavior seem apparent, as can be seen from plotting multiple weeks alongside, as in Figure \ref{fig:weeksPlot}.

\begin{figure}[H]
```{r weeksPlot,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE}
weeksPlot
```
\caption{Attack sums from all  weeks in the experimental period plotted against week days. No pattern seems to arise.}
\label{fig:weeksPlot}
\end{figure}

We analyzed the data from two perspective: we ran a before-and after analysis, comparing the summarized aggression levels before and after the intervention period (with various additional predictors), and a time-series perspective, which took a more fine-grained perspective. For now, we will focus on  the Bayesian before-and-after analysis, for which the data were  cleaned and converted into a summarized form, involving the variables listed in Table \ref{tab:baaVars}.

\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
variable <- c("AB", "AD", "AA", "CB", "CD", "CA", "group", "IC")
explanation <- c("attacks before (pre-treatment)", "attacks during (the treatment period)",
                 "attacks after (post-treatment)", "comments before", "comments during",
                 "comments after", "treatment group", "intervention count")
vars <- data.frame(variable, explanation)
#mykable(vars)
```
\normalsize

\begin{table}
\centering
\begin{tabular}{ll}
\toprule
variable & explanation\\
\midrule
\cellcolor{gray!6}{AB} & \cellcolor{gray!6}{attacks before (pre-treatment)}\\
AD & attacks during (the treatment period)\\
\cellcolor{gray!6}{AA} & \cellcolor{gray!6}{attacks after (post-treatment)}\\
CB & comments before\\
\cellcolor{gray!6}{CD} & \cellcolor{gray!6}{comments during}\\
CA & comments after\\
\cellcolor{gray!6}{group} & \cellcolor{gray!6}{treatment group}\\
IC & intervention count\\
\bottomrule
\end{tabular}

\caption{Variables involved in the before-and-after analysis.}
\label{tab:baaVars}
\end{table}

Further variables were defined in terms of those, in particular, we will be predicting \textsf{AdiffS} which is the standardized difference  \textsf{AA}-\textsf{AB}, and \textsf{AdiffS}, which is the standardized difference \textsf{CA}-\textsf{CB}. The predictors were also standardized (and named $\langle$variable\char`_name$\rangle$S),  and  a numerical index for the group (\textsf{groupID})  has been introduced.

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
summaries$ABS <- standardize(summaries$AB)
summaries$CBS <- standardize(summaries$CB)
summaries$AAS <- standardize(summaries$AA)
summaries$CAS <- standardize(summaries$CA)
summaries$CDS <- standardize(summaries$CD)
summaries$ADS <- standardize(summaries$AD)
summaries$group <- as.factor(summaries$group)
summaries$groupID <-  as.integer( as.factor(summaries$group) )
```
\normalsize

The distribution of \textsf{IC} in the treatment groups is visualized in Figure \ref{fig:interventionsDistro}. Note that the distributions are somewhat different, even though the total intervention counts are similar (`r  colSums(table(summaries$IC, summaries$group))["empathy"]` for empathy and  `r colSums(table(summaries$IC, summaries$group))["normative"]` for normative). The issue is discussed in Section XXXXX.

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
interventionsDistro <- ggplot(summaries[summaries$group != "control",], aes(x = IC, fill = group))+
  geom_bar()+theme_tufte()+
  xlab("interventions received")+
  labs(title = "Intervention counts in treatment groups")+
  scale_x_continuous(breaks = seq(0,40,5))
```
\normalsize

\begin{figure}[H]
```{r interventionsDistro,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE}
interventionsDistro
```
\caption{Distribution of daily interventions, by treatment group.}
\label{fig:interventionsDistro}
\end{figure}

Second, when we look at the distribution of standardized difference in attacks, when restricted to (-1,1), the peaks of distributions are shifted a bit between the groups, with lowest median for the normative group, but the differences seem minor (Figure \ref{fig:violJoint}).  This might suggest no impact of the interventions, but this conclusion would be too hasty, as the impact of  other predictor variables and interactions involved can mask actual associations. We will take a closer look at this issue in our analysis.

\vspace{1mm}
\footnotesize
```{r violEmpiricalAdiff,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE}
violAdiffS <- ggplot(summaries, aes(x=group, y = AdiffS))+
  geom_violin() +theme_tufte() +theme(plot.title.position = "plot")
violJoint <- ggarrange(violAdiffS+ggtitle("whole range"),
              violAdiffS + ylim(c(-1,1))+geom_boxplot(width = .2)+
              ggtitle("restricted to (-1,1)"))   +theme(plot.title.position = "plot")
violJointTitled <- annotate_figure(violJoint, 
  top = text_grob("Empirical distribution of change in attacks (standardized)",
                  size = 12))
```
\normalsize

\begin{figure}[H]
```{r violJoint,echo=FALSE,eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "100%",   message = FALSE, warning = FALSE, results = FALSE}
violJointTitled
```
\caption{Empirical distribution of change in attacks, by treatment group.}
\label{fig:violJoint}
\end{figure}

To see how this masking can occur, let's inspect  changes in attacks against intervention counts. It turns out that  restricting attention to various activity levels results in fairly strong changes to  the regression lines (Figure \ref{fig:linearShift}).

\vspace{1mm}
\footnotesize
```{r ic,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE}
icplot1 <- ggplot(summaries, aes(x = IC, y = AdiffS, color = group, fill = group))+
  geom_jitter(alpha = 0.6, size =.8)+theme_tufte()+ theme(plot.title.position = "plot") +
  geom_smooth(alpha = 0.2, method = "lm")+
  xlim(c(0,25))+ylim(c(-2,2))+
  ggtitle("sd restricted to (-2,2)")+
  theme(legend.position = c(0.65, 0.2))

icplot2 <-  ggplot(summaries, aes(x = IC, y = AdiffS, color = group, fill = group))+
  geom_jitter(alpha = 0.6, size =.8)+theme_tufte()+theme(plot.title.position = "plot") +
  geom_smooth(alpha = 0.2, method = "lm")+
  xlim(c(0,25))+ylim(c(-1,1))+ggtitle("sd restricted to (-1,1)")+
  theme(legend.position = c(0.65, 0.2))

icplotJoint <- ggarrange(icplot1, icplot2) 
icplotTitled <- annotate_figure(icplotJoint, 
  top = text_grob("Change in attacks (standardized) vs interventions received",  size = 12))
```
\normalsize

\begin{figure}
```{r fig:linearShift,echo=FALSE,eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "100%",   message = FALSE, warning = FALSE, results = FALSE}
icplotTitled  
```
\caption{Change in attacks vs intervention counts by treatment group, jittered with  linear smoothing.}
\label{fig:linearShift}
\end{figure}

Some interactions are also suggested by the differences in linear smoothing when attention is restricted when it comes to change in comments.

<!-- \vspace{1mm} -->
<!-- \footnotesize -->
<!-- ```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"} -->
<!-- #head(summaries) %>% kable( "latex", booktabs = T) %>%  -->
<!-- #  kable_styling(latex_options = c("striped", "scale_down") ,font_size = 9) -->

<!-- ``` -->
<!-- \normalsize -->